\addcontentsline{toc}{chapter}{Acknowledgments}
\topskip0pt
\vspace*{2cm}
\Huge
\textbf{Acknowledgments} \label{sec:acknowledgments}
\normalsize

\vspace{1cm}

These lines end a five year endeavour, during which I have met people, and learnt things that will stick with me for the years to come.
Many have given me a hand along the way, but there are a certain few without which this work would have never seen the light.
I would like to take this moment and thank them for their help.

First of all, I would like to thank Ricard Delgado and the Swiss Center for Electronics and Microtechnology for hosting me during the six month internship that enabled me to focus completely on my work.
Ricard has been an unsurpassable host and, together with Enric Muntan\'e, they made Neuch\^atel feel like a home away from home.
Secondly, I would like to thank Valerio Schiavoni from the University of Neuch\^atel.
It was him who proposed the initial topic, it was him who provided the contact with Peter Pietzuch and Pierre-Louis Aublin, who gladly granted me with developer access to their on-going projects, and it was him who, week in and week out, helped me keep the bigger picture and paved the way for the personal success this work has been.
Lastly, I would like to thank Professor Jos\'e Adri\'an Rodr\'iguez Fonollosa for always answering my doubts and helping me without hesitation.

On a more personal note, I would like to thank my parents for their unconditional love and support throughout all these years.
Without their guidance during decisive moments, I would not be writing these lines today.
I would also like to express my most sincere gratitude to Jos\'e Andr\'es, only he knows what has taken us to be here today, and I owe a good part of it to him.
Finally, I must devote some words to Clara.
Even though time might now be scarce, she has taught me more about myself than I would have learnt in a lifetime without her.

\vspace{1cm}

\begin{flushright}
Carlos Segarra Gonz\'alez

Neuch\^atel, \today
\end{flushright}

\vspace*{\fill}
