\addcontentsline{toc}{chapter}{Note from the Author}
\topskip0pt
\vspace*{4cm}
\Huge
\textbf{Note from the Author} \label{sec:acknowledgments}
\normalsize

\vspace{1cm}

The work here presented is my Bachelor's Thesis for my joint degree in Mathematics and Telecommunications Engineering within the Interdisciplinary Higher Education Center (CFIS) from the Polytechnic University of Catalonia (UPC).
It has been developed during a six-month internship at the Swiss Center for Electronics and Microtechnology (CSEM) under the supervision of Ricard Delgado-Gonzalo, in the Embedded Software group.
Professor Jos\'e Adri\'an Rodr\'iguez Fonollosa, from the UPC, has co-advised this project, and has been the official tutor with regard to the university.

Parts of this work have been included in two different conference papers which have been accepted and will be published in their respective proceedings.
The first one is entitled \textbf{\textit{Secure Stream Processing for Medical Data}}~\cite{Segarra2019} and will be presented at the 41st IEEE Engineering in Medicine and Biology Conference (EMBC '19) to be held in Berlin, Germany, from July 23-27 2019.
It focuses on a particular medical application our work could be used in.
The second one is entitled \textbf{\textit{Using Trusted Execution Environments for Secure Stream Processing of Medical Data}}~\cite{Segarra2019b} and will be presented in the 19th International Conference on Distributed Applications and Interoperable Systems (DAIS '19) to be held in Copenhagen, Denmark, from June 17-21 2019.
It covers our solution's implementation in-depth and evaluates its performance.

%\vspace{1cm}
%
%\begin{flushright}
%Carlos Segarra Gonz\'alez
%
%Neuch\^atel, \today
%\end{flushright}

\vspace*{\fill}
